Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Document Preparation Systems

### 2018-11-16 main.tex
Added LaTeXML explanations (close #8)

\usepackage[utf8]{inputenc}
\usepackage[english]{babel} \usepackage[T1]{fontenc} 
\usepackage{hyperref}			
\usepackage[acronym, nomain]{glossaries} \usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{parskip}		
\usepackage{txfonts}            
\usepackage{xcolor}				
\usepackage{enumitem}			
\usepackage{algorithm}
\usepackage{algpseudocode} 
\usepackage{lipsum} 
\usepackage{tabularx}
\usepackage{stmaryrd}
\usepackage{float}
\usepackage{csquotes}
\usepackage{layout}
\usepackage{textcomp} 
\usepackage{authblk}

@@

@@

\section{Conclusion \& Future Work}\label{ch:conc-future-work}
%The results and conclusions of the translations of the formula set from the \DLMF{} are diverse. Some test cases have shown us that the set of semantic macros are not perfect yet and need to be improved. This is seen in the example of the macro definition for the Wronskian determinant. However, simply adding more and more macros to improve the set is not the right decision. In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use. If the system is no longer handy, it will not be used {\bf [HSC: I don't agree with this paragraph at all. In fact, I think we've had this discussion already in the past and it was resolved. It needs to be resolved again. There are many examples where enhancing macros is a benefit. Not only that, we've discussed this recently in our publications. This needs to be fixed prior to submission. I vote for removing this paragraph altogether until we can resolve this issue.]} {\bf [AGP: Since the paper in it's current state is quite long and the section works well without this paragraph, I just commented it out and avoid the discussion here...]}

The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia. The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions. Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}. However, combining those techniques and automatically running translation checks, not only can discover errors in mathematical compendia but can also detect errors in simplifications or computations of the \gls*{cas}.

@@

\emergencystretch=1em
\printbibliography
\end{document}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[english]{babel} 
\usepackage[T1]{fontenc}
\usepackage{hyperref}			
\usepackage[acronym, nomain]{glossaries} 
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{parskip}		
\usepackage{txfonts}			
\usepackage{enumitem}			
\usepackage{algorithm}
\usepackage{algpseudocode} 
\usepackage{lipsum} 
\usepackage{tabularx}
\usepackage{stmaryrd}
\usepackage{float}
\usepackage{csquotes}
\usepackage{layout}
\usepackage{textcomp} 
\usepackage{authblk}

\usepackage[dvipsnames]{xcolor}
\usepackage{lstomdoc}
\lstset{language=MathML,%basicstyle=\sf\footnotesize,
	columns=fullflexible,
    belowcaptionskip=0em,
    belowskip=-1.5em,
    alsoletter={'},
    stringstyle = {\color{Blue}},
    keywordstyle = {\color{OliveGreen!70!LimeGreen}},
    keywordstyle = [2]{\color{Fuchsia}},
    morekeywords = [2]{cd,definitionURL,id,xmlns}
}

@@

Miller also developed LaTeXML, a tool for converting \LaTeX{} expressions to MathML~\parencite{LaTeXML}. LaTeXML was used to generate the \gls*{dlmf} and is able to parse the \Macro s to content MathML. Even though many \gls*{cas} are able to import and export MathML, they fail for special functions. Schubotz et.~al recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.\Cref{sec:trans-mml} also shows three examples of generated MathML by LaTeXML, \Maple, and \Mathematica. We did not use MathML as an intermediate format to perform translations between \LaTeX{} and \gls*{cas}, because of the problems content MathML has~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.

@@

\section{Conclusion \& Future Work}\label{ch:conc-future-work}
The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia. The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions. Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}. However, combining those techniques and automatically running translation checks, not only can discover errors in mathematical compendia but can also detect errors in simplifications or computations of the \gls*{cas}.

@@

\emergencystretch=1em
\printbibliography

\appendix
\section*{Appendix}
\section{Translations via MathML}\label{sec:trans-mml}
The problems to use MathML data as an intermediate format can be demonstrated by using LaTeXML, \Maple, and \Mathematica{} to generate content MathML from their inputs. While the inputs for \gls*{cas} are unique, ideally all systems should produce the same content MathML data. Also, for semantic \LaTeX, LaTeXML should be able to generate the correct content MathML data. As an example we generated content MathML via LaTeXML for
\begin{equation}\label{eq:jac-lml}
\verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}|,
\end{equation}
and via \Maple{} and \Mathematica{} for the manually verified appropriated  translations of (\ref{eq:jac-lml})
\begin{eqnarray}
&& \verb|JacobiP(n,alpha,beta,cos(a*Theta))|,\text{ and}\label{eq:jac-map}\\
&& \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|.\label{eq:jac-mat}
\end{eqnarray}
Listings~\ref{lst:latexml}-~\ref{lst:mathematica} shows the generated content MathML. As the listings show, all generated MathML distinguish to each other. The problems are not only flaws in their export implementations but also in the fact that even content MathML is not unique. Based on this fact and the differences in the generated content MathML between each system, we decided to not use MathML as an intermediate format.

\begin{lstlisting}[label={lst:latexml},mathescape=true,caption=The generated content MathML for~(\ref{eq:jac-lml}) using LaTeXML with the command \texttt{latexmlc} and the set of semantic macros.]
<math xmlns="http://www.w3.org/1998/Math/MathML" $\ldots$>
  <semantics>
    <apply>
      <apply>
        <csymbol cd="dlmf">Jacobi-polynomial-P</csymbol>
        <ci>$\alpha$</ci>
        <ci>$\beta$</ci>
        <ci>$n$</ci>
      </apply>
      <ci>$x$</ci>
    </apply>
  </semantics>
</math>
\end{lstlisting}

\vspace{0.4cm}
\begin{lstlisting}[label={lst:maple},mathescape=true,caption=The content MathML for~(\ref{eq:jac-map}) generated by Maple 16 (including manually added linebreaks). We were using the \texttt{MathML[ExportContent]} command to produce the output.]
<math xmlns='http://www.w3.org/1998/Math/MathML'>
  <apply id='id6'>
    <csymbol id='id1' 
      definitionURL='http://www.maplesoft.com/MathML/JacobiP'>
        JacobiP
    </csymbol>
    <ci id='id2'>n</ci>
    <ci id='id3'>alpha</ci>
    <ci id='id4'>beta</ci>
    <ci id='id5'>x</ci>
  </apply>
</math>
\end{lstlisting}

\vspace{0.4cm}
\begin{lstlisting}[label={lst:mathematica},mathescape=true,caption=The content MathML for~(\ref{eq:jac-mat}) generated by Mathematica. We were using the \texttt{ExportString} command to produce the output.]
<math xmlns='http://www.w3.org/1998/Math/MathML'>
 <apply>
  <ci>JacobiP</ci>
  <ci>n</ci>
  <ci>&#945;</ci>
  <ci>&#946;</ci>
  <ci>x</ci>
 </apply>
</math>
\end{lstlisting}
\end{document}
